Tree establishment in models of forest dynamics: a key priority for further research

### 2023-03-12 manuscript/manuscriptRW.Rmd
changes in lats paragraph abstract

Using 15 models and model variants built to capture long-term (i.e.,
decadal to centennial) forest dynamics at the stand, landscape and
global levels, we simulate tree establishment at 200 sites representing
environmental gradients across Europe. The results are benchmarked
against comprehensive data from unmanaged forests.

@@

Even though the overestimation of tree establishment represents a
challenge for the usefulness of model projections under uncertain
conditions such as climate change, and although some formulations should
be reconsidered to correct excessive ingrowth levels, it is remarkable
that most of the models capture the essential features of the highly
complex process of tree establishment, while not having been
parameterized with such data. We conclude that in spite of all the
uncertainties, a lot can be gained by a focus on the modeling of
establishment processes in these model, this will make their long-term
forest dynamics projections under changing environmental conditions much
more robust.

>>>>

Using 15 models and model variants built to capture long-term (i.e.,
decadal to centennial) forest dynamics at the stand, landscape and
global levels, we simulate tree establishment at 200 sites representing
environmental gradients across Europe. The results are evaluated against
comprehensive data from unmanaged forests.

@@

In spite of more or less substantial mismatches between model results
and data, it is remarkable that most of the models capture the essential
features of the highly complex process of tree establishment, while not
having been parameterized with such data. We conclude that in spite of
all the uncertainties, a lot can be gained by a focus on the modeling of
establishment processes in these model, this will make their long-term
forest dynamics projections under changing environmental conditions much
more robust.
